---
title: "ChiRP: Chinese Restaurant Process Models for Regression and Clustering"
author: "Arman Oganisian"
date: "2/2/2019"
output:
  html_document:
    includes:
      in_header: "favicon.html" 
---

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"logo.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation 
`ChiRP` is an R package that implements **Chi**nese **R**estaurant **P**rocess mixtures of regression models. The package currently supports zero-inflated continuous outcomes, continous outcomes, and binary outcomes. Install from GitHub as follows:

```{r install, include=T, eval=F}
## install.packages('devtools' ) ## make sure to have devtools intalled 
devtools::install_github('stablemarkets/ChiRP')
library(ChiRP)
```

## Contact and Issue Reporting
Please [report issues](https://github.com/stablemarkets/ChiRP/issues) on GitHub. Contact via Email ( [aoganisi@upenn.edu](aoganisi@upenn.edu) ) or Twitter ( [@StableMarkets](https://twitter.com/StableMarkets)  ). 

## References



