---
title: "ChiRP: Chinese Restaurant Process Models for Regression and Clustering"
author: "Arman Oganisian"
date: "2/2/2019"
output:
  html_document:
    includes:
      in_header: "favicon.html" 
---

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"logo.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation 
`ChiRP` is an R package that implements **Chi**nese **R**estaurant **P**rocess mixtures models for regression and clustering. The package currently supports zero-inflated continuous outcomes, continous outcomes, and binary outcomes. 

Install from GitHub as follows:
```{r install, include=T, eval=F}
## install.packages('devtools' ) ## make sure to have devtools intalled 
devtools::install_github('stablemarkets/ChiRP')
library(ChiRP)
```

## What Are Chinese Restaurant Process Models?
CRP models, aka Dirichlet Process (DP) models, are a class of Bayesian nonparametric models. They provide a very flexible fit to complex data while also providing easy uncertainty estimates via posterior inference. They work by partitioning complex data into more homogenous clusters and modeling each with a locally parametric model. But don't be fooled! While they use locally parametric models, CRP models assume there are infinitely many clusters so that the parameter space is infinite-dimensional - making this a truly nonparmatric method. Please see the examples page for uses.

## How to Cite this Package
Since this package was written in conjunction with [this paper](https://arxiv.org/abs/1810.09494), please cite it when using this package. Thank you!

## Contact and Issue Reporting
Please [report issues](https://github.com/stablemarkets/ChiRP/issues) on GitHub. 

Package author and maintainer: Arman Oganisian.
Contact via Email ( [aoganisi@upenn.edu](aoganisi@upenn.edu) ) or  [Twitter](https://twitter.com/StableMarkets). 

## Thanks and Acknowledgements
Thanks to [Jason Roy](https://twitter.com/jasonroy) for invaluable discussions regarding underlying MCMC computations. Special thanks to [Nick Illenberger](https://twitter.com/_Willy_Nilly) and [Caroyln Lou](https://twitter.com/carelouu) for designing the package hex and coming up with the creative name, `ChiRP`!

## References
[A Bayesian Nonparametric Method for Estimating Causal Treatment Effects on Zero-Inflated Outcomes](https://arxiv.org/abs/1810.09494). A. Oganisian et al. 2018.

[Bayesian nonparametric generative models for causal inference with missing at random covariates](https://onlinelibrary.wiley.com/doi/abs/10.1111/biom.12875). Roy et al. 2018.

[A Bayesian nonparametric approach to marginal structural models for point treatments and a continuous or survival outcome](https://www.ncbi.nlm.nih.gov/pubmed/27345532). Roy et al. 2016.


[Dirichlet Process Mixtures of Generalized Linear Models](http://www.cs.columbia.edu/~blei/papers/HannahBleiPowell2011.pdf). Hannah et al. 2011.

[Markov Chain Sampling Methods for Dirichlet Process Mixture Models](http://www.stat.columbia.edu/npbayes/papers/neal_sampling.pdf). Radford Neal. 2000.


