---
title: "ChiRP: Chinese Restaurant Process Models for Regression and Clustering"
author: "Arman Oganisian"
output:
  html_document:
    includes:
      in_header: "favicon.html" 
---

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"logo.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation 
`ChiRP` is an R package that implements **Chi**nese **R**estaurant **P**rocess mixtures models for regression and clustering. The package currently supports zero-inflated continuous outcomes, continuous outcomes, and binary outcomes. 

Install from GitHub as follows:
```{r install, include=T, eval=F}
## install.packages('devtools' ) ## make sure to have devtools installed 
devtools::install_github('stablemarkets/ChiRP')
library(ChiRP)
```

Help documentation in `R` is also available.  After installing the package and loading it with `library()`, use `?` to access help documentation for specific functions:
```{r helpfiles, include=T, eval=F}
?ChiRP::NDPMix  # for continuous outcomes
?ChiRP::ZDPMix  # for zero-inflated, semi-continuous outcomes
?ChiRP::PDPMix  # for binary outcomes
?ChiRP::cluster_assign_mode # computes posterior mode cluster assignment
```

## What Are Chinese Restaurant Process Models?
CRP models, aka Dirichlet Process (DP) models, are a class of Bayesian nonparametric models. They provide a very flexible fit to complex data while also providing easy uncertainty estimates via posterior inference. They work by partitioning complex data into more homogenous clusters and modeling each with a locally parametric model. But don't be fooled! While they use locally parametric models, CRP models assume there are infinitely many clusters so that the parameter space is infinite-dimensional - making this a truly nonparametric method. Please see the examples page for uses.

## How to Cite this Package
Since this package was written in conjunction with [this paper](https://arxiv.org/abs/1810.09494), please cite it when using this package. BibTeX

The software itself was [published](https://joss.theoj.org/papers/3b83a0a3f1220f97657a1075b78e480a) in the Journal of Open Source Software.

```{r bibtex, include=T, eval=F}
@article{Oganisian2019ChiRP,
	journal = {Journal of Open Source Software},
	doi = {10.21105/joss.01287},
	issn = {2475-9066},
	number = {35},
	publisher = {The Open Journal},
	title = {ChiRP: Chinese Restaurant Process Mixtures for Regression and Clustering},
	url = {http://dx.doi.org/10.21105/joss.01287},
	volume = {4},
	author = {Oganisian, Arman},
	pages = {1287},
	date = {2019-03-26},
	year = {2019},
	month = {3},
	day = {26},
}
```

## Contact, GitHub, and Issue Reporting
The package is stored on [GitHub](https://github.com/stablemarkets/ChiRP), where you can [report issues](https://github.com/stablemarkets/ChiRP/issues).

Development Status:

[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Build Status](https://travis-ci.org/stablemarkets/ChiRP.svg?branch=master)](https://travis-ci.org/stablemarkets/ChiRP)
[![Coveralls github](https://img.shields.io/coveralls/github/stablemarkets/ChiRP.svg?style=popout)](https://coveralls.io/github/stablemarkets/ChiRP)


Package author and maintainer: Arman Oganisian.
Contact via Email ( [aoganisi@upenn.edu](aoganisi@upenn.edu) ) or  [Twitter](https://twitter.com/StableMarkets). 

## Contributing to `ChiRP`
You can contribute in two ways:

1. Contribute to base code: First, start an issue in this repository with the proposed modification. Fork this repository, make changes/enhancements, then submit a pull request. The issue will be closed once the pull request is merged.
2. Contribute an example: First, start an issue in the [companion site's repository](https://github.com/stablemarkets/ChiRPsite). Fork the repository and add a new example to `examples.Rmd`. Use `rmarkdown::render_site()` to build the site. Submit a pull request in that same repository. The issue will be closed once updates are merged.


## Acknowledgements
Thanks to [Jason Roy](https://twitter.com/jasonroy) for invaluable discussions regarding underlying MCMC computations. Special thanks to [Nick Illenberger](https://twitter.com/_Willy_Nilly) and [Caroyln Lou](https://twitter.com/carelouu) for designing the package hex and coming up with the creative name, `ChiRP`!

This work was supported in part by Grant R01GM112327 from National Institute Of General Medical
Sciences.

## References
[A Bayesian Nonparametric Method for Estimating Causal Treatment Effects on Zero-Inflated Outcomes](https://arxiv.org/abs/1810.09494). A. Oganisian et al. 2018.

[Bayesian nonparametric generative models for causal inference with missing at random covariates](https://onlinelibrary.wiley.com/doi/abs/10.1111/biom.12875). Roy et al. 2018.

[A Bayesian nonparametric approach to marginal structural models for point treatments and a continuous or survival outcome](https://www.ncbi.nlm.nih.gov/pubmed/27345532). Roy et al. 2016.


[Dirichlet Process Mixtures of Generalized Linear Models](http://www.cs.columbia.edu/~blei/papers/HannahBleiPowell2011.pdf). Hannah et al. 2011.

[Markov Chain Sampling Methods for Dirichlet Process Mixture Models](http://www.stat.columbia.edu/npbayes/papers/neal_sampling.pdf). Radford Neal. 2000.

